\documentclass[a4paper, twoside, colorlinks=true, allcolors=blue]{article}
%% Language and font encodings
\usepackage[%
style=phys, %
articletitle = false, biblabel = brackets, %
pageranges = false, chaptertitle = false, url = false]{biblatex}
\AtBeginBibliography{%
\urlstyle{rm}%
}
\addbibresource{bibs/ref.bib}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsthm, dsfont}
\usepackage{mathtools}
\usepackage{bbold}
\usepackage{physics}
\usepackage{pgfplots}
\usepackage{mmacells}
\usepackage{changepage}
\pgfplotsset{compat = 1.8}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{eufrak}
%% Useful packages
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[labelfont={bf,footnotesize}]{caption}
\usepackage{subcaption}


\usepackage[colorinlistoftodos]{todonotes}
\usepackage[]{hyperref}
\numberwithin{equation}{section}
\DeclareCiteCommand{\supercite}[\mkbibsuperscript]
  {\iffieldundef{prenote}
     {}
     {\BibliographyWarning{Ignoring prenote argument}}%
   \iffieldundef{postnote}
     {}
     {\BibliographyWarning{Ignoring postnote argument}}%
   \bibopenbracket}%
  {\usebibmacro{citeindex}%
   \usebibmacro{cite}}
  {\supercitedelim}
  {\bibclosebracket}

\newcommand{\dbar}{d\hspace*{-0.08em}\bar{}\hspace*{0.1em}}
\newcommand{\bQ}{\mathds{Q}}
\newcommand{\bR}{\mathds{R}}
\newcommand{\bC}{\mathds{C}}
\newcommand{\bN}{\mathds{N}}
\newcommand{\bZ}{\mathds{Z}}
\newcommand{\bP}{\mathds{P}}
\newcommand{\bE}{\mathds{E}}
\newcommand{\rmd}{\mathrm{d}}
\newcommand{\entpp}{\dot{\mathcal{S}}}
\newcommand{\Overline}[1]{\overline{\overline{#1}}}
\newcommand{\ai}[0]{\scalebox{0.7}{$\square$}}

\setlength\parindent{0pt}
\setlength\parskip{5pt}

% \DeclarePairedDelimiter\abs{\lvert}{\rvert}
% \makeatletter
% \let\oldabs\abs
% \def\abs{\@ifstar{\oldabs}{\oldabs*}}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem*{prop}{Proposition}


\title{Path Probabilities and Entropy Production of Non-Markov Processes}
\author{Farid Kaveh}
% Update supervisor and other title stuff in title/title.tex

\begin{document}
\input{title/title.tex}
\renewcommand{\abstractname}{\normalsize Declaration}
\begin{abstract}\normalsize
    The work contained herein is my own unless otherwise indicated. 
    
    
    
    \vspace{1cm}
    \quad \quad Farid Kaveh
\end{abstract}
\newpage
\renewcommand{\abstractname}{\normalsize Abstract}
\begin{abstract}\normalsize 
We investigate the entropy production rate of several concrete examples of non-Markov systems using a path-space approach. The validity of this approach is confirmed through application to the telegraph process. An exact closed-form expression for the entropy production rate of a current-free system is presented. We develop a perturbation theory for the entropy production rate of one-dimensional diffusion processes with stochastic, time-dependent drift coefficient when the process driving the drift is not experimentally accessible. This theory is applied to an asymmetric Run-and-Tumble particle to derive the leading order contribution to the entropy production rate of this process. We conclude by identifying some unifying paradigms among `non-Markov' processes that have attracted attention in the stochastic thermodynamics literature and relating back to the theory of stochastic processes. 
\end{abstract}
\newpage
\renewcommand{\abstractname}{\normalsize Acknowledgements}
\begin{abstract}\normalsize
I would like to thank everyone in the Non-Equilibrium Systems group at Imperial College London. I learned much from this group and our daily conversations were a constant source of inspiration for me. Special acknowledgments go to

\begin{itemize}
\item my supervisor, Dr Gunnar Pruessner, for his indispensable insight and guidance, 
\item Jacob Knight, whose contributions to this text cannot be overstated, and 
\item Luca Cocconi and Connor Roberts, who provided much needed criticism of this thesis. 
\end{itemize}
\end{abstract}
\newpage
\renewcommand{\abstractname}{\normalsize Dedication}
\begin{abstract}\normalsize
\vspace{3cm}
\textit{``All is number''} 
\vspace{2cm}


\begin{quote}
``No one knows what entropy
really is, so in a debate you will always
have the advantage.''
\end{quote}

\hfill ---Jon von Neumann to Claude E. Shannon.
\end{abstract}
\newpage

\tableofcontents
\listoffigures
\newpage
%\input{introduction/introduction.tex}
%\input{background/background.tex}
%\input{project/project.tex}
%\input{evaluation/evaluation.tex}
%\input{conclusion/conclusion.tex}
%\input{appendix/appendix.tex}
\input{introduction/intro.tex}
\input{two_state_system/two_state_system.tex}
\input{the_waiting_room/waiting_room.tex}
\input{Unreq_love/UL.tex}
\input{asymmetric_RnT/asymmetric_Rnt.tex}
\input{classification/classification.tex}
\input{conclusion/conclusion.tex}
\input{Appendix/appendix.tex}




\printbibliography

\end{document}
